Heuristics and Really Hard Instances for Subgraph Isomorphism Problems

### 2016-02-02 paper/paper.tex
Final typos

\author{First Author \thanks{This work was supported by a small pile of money and a large
    electricity bill.} \and Second Author
    \and Third Author \\
Some Instituion, Somewhere, Some Country \\
anonymous@example.com}

@@

We might guess that degree-based heuristics would just not work for the induced problem: for any
claim about the degree, the opposite will hold for the complement constraints. However, empirically,
this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
(The only steps performed by the Glagsow algorithm which differs under taking the complements are
the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
step using degree information, but does not consider the complement degree, and VF2 uses
connectivity in the pattern graph.)

@@

In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
heuristics \citep{Carletti:2015}. In this situation, these experiments justify re-using the
non-induced heuristics on induced problems.

@@

The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target
vertex $w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to
$w$.  Constraints are added to ensure that each pattern vertex maps to exactly one target vertex,
that each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are
mapped to adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We
used the Clasp solver~\citep{gekakaosscsc11a} version 3.1.3 to solve the pseudo-boolean instances.
The instances that are hard for the Glasgow solver remain hard for the PB solver, including
instances inside the central region, and the easy satisfiable instances remain easy.

We also implemented (but do not plot) an equivalent SAT encoding with direct encoding of the
cardinality constraints.  Using the Glucose solver, we saw very similar performance to the PB
encoding. We also implemented a similar integer program encoding; the Gurobi solver was only able to
solve some of the trivial satisfiable instances, and was almost never able to prove unsatisfiability
within the time limit.

@@

These results suggest that the central region may be genuinely hard, despite not being near a phase
transition. The clique encoding in particular rules out the hypothesis that subgraph isomorphism
solvers only find this region hard due to not reasoning simultaneously about adjacency and
non-adjacency, since the association graph encoding constraints consider compatibility rather than
adjacency and non-adjacency.

>>>>

\author{First Author \thanks{This work was supported by a small pile of money from an anonymous
        funding agency, and a large electricity bill.} \and Second Author
    \and Third Author \\
Some Instituion, Somewhere, Some Country \\
anonymous@example.com}

@@

We might guess that degree-based heuristics would just not work for the induced problem: for any
claim about the degree, the opposite will hold for the complement constraints. However, empirically,
this is not the case: on the final row of \cref{figure:induced}, we show whether it is better to use
the original pattern and target as the input to the Glasgow algorithm, or to take the complements.
(The only steps performed by the Glagsow algorithm which differ under taking the complements are
the degree-based heuristics.  LAD and VF2 are not symmetric in this way: LAD performs a filtering
step using degree information, but does not consider the complement degree, and VF2 uses
connectivity in the pattern graph.)

@@

In practice, this is unlikely to be a problem: most real-world instances are extremely sparse and
are usually easy, which perhaps explains the continuing popularity of VF2's connectivity-based
heuristics \citep{Carletti:2015}. In this situation, these experiments justify reusing the
non-induced heuristics on induced problems.

@@

The pseudo-boolean (PB) encoding is as follows. For each pattern vertex $v$ and each target vertex
$w$, we have a binary variable which takes the value 1 if and only if $v$ is mapped to $w$.
Constraints are added to ensure that each pattern vertex maps to exactly one target vertex, that
each target vertex is mapped to by at most one pattern vertex, that adjacent vertices are mapped to
adjacent vertices, and that non-adjacent vertices are mapped to non-adjacent vertices. We used the
Clasp solver~\citep{gekakaosscsc11a} version 3.1.3 to solve the pseudo-boolean instances.  The
instances that are hard for the Glasgow solver remain hard for the PB solver, including instances
inside the central region, and the easy satisfiable instances remain easy. (Similar results were
seen with the Glucose SAT solver using a direct encoding of the cardinality constraints. We also
implemented an integer program encoding; the Gurobi solver was only able to solve some of the
trivial satisfiable instances, and was almost never able to prove unsatisfiability within the time
limit.)

@@

These results suggest that the central region may be genuinely hard, despite not being near a phase
transition. The clique results in particular rule out the hypothesis that subgraph isomorphism
solvers only find this region hard due to not reasoning simultaneously about adjacency and
non-adjacency, since the association graph encoding constraints consider compatibility rather than
adjacency and non-adjacency.
